---
title: "Glasso"
author:
- name: Elyas Heidari
  affiliation: 
  - &IMLS Institute for Molecular Life Sciences, University of Zurich, Switzerland
  - Swiss Institute of Bioinformatics (SIB), University of Zurich, Switzerland
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: false
---

# Setup / definitions

## Libraries

```{r setup_knitr, include=FALSE}
library('BiocStyle')
options(bitmapType='cairo')
knitr::opts_chunk$set(autodep=TRUE, cache=FALSE, dev='png', cache.lazy = FALSE)
knitr::opts_knit$set( root.dir='..' )
set.seed(1996)
# wflow_build(files='analysis/ms10_glasso_vignette.Rmd', view=F, verbose=T, delete_cache=T)
```

## Helper functions
```{r setup_libs, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
suppressMessages(source('code/scGCN/dep.R')) 
source('code/ms00_utils.R')
source('code/ms10_glasso.R')
```

## Inputs
```{r setup_input, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
# results files
base_dir = '/home/Shared_taupo/will/projects/MS_lesions'
sce_f    = file.path(base_dir, 'data/sce_raw/ms_sce.rds')
conos_f = file.path(base_dir, 'output/ms06_sccaf/conos_merged_dt_2021-03-18.csv')
meta_f   = file.path(base_dir, 'data/metadata/metadata_updated_20201127.txt')
pb_f     = file.path('.', 'output/ms09_muscat/sce_pb_broad_2021-03-23.rds')
```


## Outputs
```{r setup_outputs, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
# output paths
time_stamp = format(Sys.Date(), '%Y_%m_%d')
out_dir    = 'output/ms10_glasso'
dir.create(out_dir)
## parameters
set.seed(1996)
cluster_var = 'conos_merge'
prop_train  = .5
max_cells   = 10000
min_genes   = 1000
min_counts  = 100
rho         = 0.5
n_hvgs_all  = 1000
n_hvgs_comb = 500
n_hvgs_type = 500
# dirs
### output
graph_com_f = file.path(out_dir, sprintf('%s/graph_comm_%s_%s.pdf', '%s', '%s', time_stamp))
graph_act_f = file.path(out_dir, sprintf('%s/graph_act_%s_%s.png', '%s', '%s', time_stamp))
```

## Load data and preprocess
```{r load_data, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
# load data
## conos clusters + sub-sampling
conos_dt = read.csv(conos_f) %>% 
    as.data.table 
broad_freqs = rowSums(table(conos_dt$type_broad, conos_dt$conos_merge) > 0)
conos_dt$broad_freq = broad_freqs[conos_dt$type_broad]

## sce object
sce      = readRDS(sce_f) %>% .[, conos_dt$cell_id]
col_dt = as.data.table(colData(sce), keep.rownames = TRUE) %>% 
    setnames('rn', 'cell_id') 
sex_bal = min(table(col_dt$sex))
col_dt %<>% 
    .[, .SD[sample(.N, min(.N, ..sex_bal), replace=FALSE)], by=sex]
conos_dt %<>% .[col_dt, on='cell_id'] %>%
    .[, condition  := ifelse(lesion_type == 'GM' | lesion_type == 'WM', 'CTR', 'MS')] %>%
    .[, lesion_type:= factor(lesion_type, ordered=TRUE, levels=lesion_ord)] %>%
    .[, type_broad := factor(type_broad, ordered=TRUE, levels=broad_ord)] %>%
    .[, conos_merge:= factor(conos_merge)] %>%
    .[, subset_size := ..max_cells %/% unique(broad_freq), by=cluster_var] %>%
    .[, .SD[sample(.N, min(.N, subset_size), replace=FALSE)], by=cluster_var]
sce  %<>%  .[,conos_dt$cell_id] %>%
    .[rowSums(counts(.)) > min_counts, ]
rowData(sce)$symbol = strsplit2(rownames(sce),split='_')[,1]
sce %<>% .[-which((rowData(.)$symbol %in% ex_genes | grepl('^MT', rowData(.)$symbol))),]
log_counts  = logNormCounts(sce) %>% logcounts 

## psudo-bulk object
# pb      = readRDS(pb_f)
# assays(pb)[['soup']] = NULL
# pb_meta = colData(pb) %>%
#     as.data.table %>%
#     setnames('group_id', 'lesion_type') %>%
#     .[, lesion_type:= factor(lesion_type, ordered=TRUE, levels=lesion_ord)] %>%
#     .[, disease_status := factor(disease_status, ordered=TRUE, levels=disease_ord)] 
# rownames(pb)   = strsplit2(rownames(pb),split='_')[,1]
# # assayNames(pb) = str_replace(assayNames(pb), '/', '--')
# pb_counts = assays(pb) %>% 
#     as.list %>%
#     map(function(x) x[, colSums(x) > 0]) %>%
#     map(~(edgeR::cpm(.x) + 1)) %>% 
#     map(log10)
# pb_counts_all = assays(pb) %>% purrr::reduce(`+`) %>% edgeR::cpm(.) %>% `+`(1) %>% log10
gc()
```

#  Preprocessing / calculations

## GGM on everything

```{r ggm_all_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
#TODO: optimize rho and threshold
graph_all = counts2graph(
    mtx = log_counts,
    rho = 0.4,
    threshold = 0.04,
    n_hvgs = n_hvgs_all
)
act_all_broad  = act_colors(graph_all$data, conos_dt$type_broad)
act_all_lesion = act_colors(graph_all$data, conos_dt$lesion_type) %>%
    .[levels(conos_dt$lesion_type), ]
gc()
```

## Normal correlation on everything

```{r corr_all_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
#TODO: optimize rho and threshold
graph_corr_0.3 = counts2graph(
    mtx = log_counts,
    rho = 0.4,
    threshold = 0.04,
    cut_off   = 0.3,
    n_hvgs = n_hvgs_all,
    corr   = TRUE
)

graph_corr_0.7 = counts2graph(
    mtx = log_counts,
    rho = 0.4,
    threshold = 0.04,
    cut_off   = 0.7,
    n_hvgs = n_hvgs_all,
    corr   = TRUE
)
gc()
```

## GGM on lesion types (robustness)
```{r ggm_lesion_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
#TODO: optimize rho and threshold
graphs_lesion = list()
ari_lesion    = list()
for(t in levels(conos_dt$lesion_type)){
    idx = which(conos_dt$lesion_type == t)
    log_counts_type = graph_all$data[idx,] 
    if(sum(colSds(log_counts_type) == 0) > 0)
        log_counts_type %<>% .[, -which(colSds(.) == 0)]
    log_counts_type %<>% as.data.table
    ggm_obj = ggm(log_counts_type, rho=0.4, threshold=0.04)
    gg = ggm_obj$graph
    wi = ggm_obj$model$wi
    isolated  = which(igraph::degree(gg) == 0)
    gg = delete.vertices(gg, isolated)
    wi %<>% .[-isolated, -isolated]
    ggm_obj$graph    = gg
    ggm_obj$model$wi = wi
    graphs_lesion[[t]] = ggm_obj
    comms  = leiden(abs(wi) > 0.04)
    comm_dt  = copy(graph_all$comm_dt) %>% setkey('GENE')
    ref_comm = comm_dt[names(V(gg))]$community
    ari_lesion[[t]] = mclust::adjustedRandIndex(ref_comm, comms)
}
gc()
```

## GGM on broad types (robustness)
```{r ggm_broad_rob_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
#TODO: optimize rho and threshold
graphs_broad_rob = list()
ari_broad_rob    = list()
for(t in levels(conos_dt$type_broad)){
    idx = which(conos_dt$type_broad == t)
    log_counts_type = graph_all$data[idx,] 
    if(sum(colSds(log_counts_type) == 0) > 0)
        log_counts_type %<>% .[, -which(colSds(.) == 0)]
    log_counts_type %<>% as.data.table
    ggm_obj = ggm(log_counts_type, rho=0.4, threshold=0.04)
    gg = ggm_obj$graph
    wi = ggm_obj$model$wi
    isolated  = which(igraph::degree(gg) == 0)
    gg = delete.vertices(gg, isolated)
    wi %<>% .[-isolated, -isolated]
    ggm_obj$graph    = gg
    ggm_obj$model$wi = wi
    graphs_broad_rob[[t]] = ggm_obj
    comms    = leiden(abs(wi) > 0.04)
    comm_dt  = copy(graph_all$comm_dt) %>% setkey('GENE')
    ref_comm = comm_dt[names(V(gg))]$community
    ari_broad_rob[[t]] = mclust::adjustedRandIndex(ref_comm, comms)
}
gc()
```

## GGM on broad types (functional annotation)
```{r ggm_broad_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
#TODO: optimize rho and threshold
graphs           = list()
act_broad_fine   = list()
act_broad_lesion = list()
for(t in levels(conos_dt$type_broad)){
    idx = which(conos_dt$type_broad == t)
    log_counts_type = log_counts[,idx]
    graphs[[t]] = counts2graph(
        mtx       = log_counts_type,
        rho       = 0.25,
        threshold = 0.025,
        n_hvgs    = n_hvgs_type
    )
    act_broad_fine[[t]]   = act_colors(graphs[[t]]$data, droplevels(conos_dt[idx,]$conos_merge))
    act_broad_lesion[[t]] = act_colors(graphs[[t]]$data, droplevels(conos_dt[idx,]$lesion_type)) %>%
        .[levels(conos_dt[idx,]$lesion_type), ]
}
gc()
```

## GGM on combined broad types (functional annotation)
```{r ggm_combined_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
#TODO: optimize rho and threshold
comb_types      = list(
    micro_immune = c('Microglia', 'Immune'),
    endo_peri    = c('Endothelial cells', 'Pericytes'),
    oligo_opc    = c('Oligodendrocytes', 'OPCs / COPs'))
graphs_comb     = list()
act_comb_fine   = list()
act_comb_lesion = list()
for(t in names(comb_types)){
    print(t)
    l = comb_types[[t]]
    idx = which(conos_dt$type_broad %in% l)
    log_counts_type = log_counts[,idx]
    graphs_comb[[t]] = counts2graph(
        mtx       = log_counts_type,
        rho       = 0.3,
        threshold = 0.03,
        n_hvgs    = n_hvgs_comb
    )
    act_comb_fine[[t]]   = act_colors(graphs_comb[[t]]$data, droplevels(conos_dt[idx,]$conos_merge))
    act_comb_lesion[[t]] = act_colors(graphs_comb[[t]]$data, droplevels(conos_dt[idx,]$lesion_type)) %>%
        .[levels(conos_dt[idx,]$lesion_type), ]
}
gc()
```

## Sample projections
```{r samples_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide', eval=FALSE}
#TODO: optimize rho and threshold
# TODO: all graph
# type broad graphs
samples_trans = list()
for(t in levels(conos_dt$type_broad)){
    trans = graphs[[t]]$trans
    data_sample = t(pb_counts[[t]][rownames(trans),])
    samples_trans[[t]] = data_sample %*% trans
}
gc()
```

# Analysis

## Corralation on everything 
```{r cor_all_graph_3, fig.height=11, fig.width=11 , collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
graph_col_comm(
    graph  = graph_corr_0.3$graph, 
    lay    = graph_corr_0.3$lay, 
    grp    = graph_corr_0.3$comm_dt$color, 
    sz     = 5, 
    title  = 'Correlation on everything, cut off = 0.3', 
    labels = graph_corr_0.3$comm_dt$GENE
)

graph_col_comm(
    graph  = graph_corr_0.7$graph, 
    lay    = graph_corr_0.7$lay, 
    grp    = graph_corr_0.7$comm_dt$color, 
    sz     = 5, 
    title  = 'Correlation on everything, cut off = 0.7', 
    labels = graph_corr_0.7$comm_dt$GENE
)
```

## GGM on everything 
```{r ggm_all_graph, fig.height=11, fig.width=11 , collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
graph_col_comm(
    graph  = graph_all$graph, 
    lay    = graph_all$lay, 
    grp    = graph_all$comm_dt$color, 
    sz     = 5, 
    title  = 'GGM on everything', 
    labels = graph_all$comm_dt$GENE
)
```

##

### GGM on lesion types (robustness){.tabset}
```{r ggm_lesion_graph, fig.height=11, fig.width=11, results='asis' , collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
comm_dt = copy(graph_all$comm_dt) %>% setkey(GENE)
for(t in levels(conos_dt$lesion_type)){
    cat('#### ', t, '\n')
        gg = graph_col_comm(
            graph  = graphs_lesion[[t]]$graph, 
            lay    = layout_nicely(graphs_lesion[[t]]$graph), 
            grp    = comm_dt[names(V(graphs_lesion[[t]]$graph))]$color, 
            sz     = 5, 
            title  = sprintf('GGM on %s, ARI = %s', t, ari_lesion[[t]]), 
            labels = comm_dt[names(V(graphs_lesion[[t]]$graph))]$GENE
        )
    cat('\n\n')

}
```

##

### GGM on broad types (robustness){.tabset}
```{r ggm_broad_rob_graph, fig.height=11, fig.width=11, results='asis' , collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
comm_dt = copy(graph_all$comm_dt) %>% setkey(GENE)
for(t in levels(conos_dt$type_broad)){
    cat('#### ', t, '\n')
        gg = graph_col_comm(
            graph  = graphs_broad_rob[[t]]$graph, 
            lay    = layout_nicely(graphs_broad_rob[[t]]$graph), 
            grp    = comm_dt[names(V(graphs_broad_rob[[t]]$graph))]$color, 
            sz     = 5, 
            title  = sprintf('GGM on %s, ARI = %s', t, ari_broad_rob[[t]]), 
            labels = comm_dt[names(V(graphs_broad_rob[[t]]$graph))]$GENE
        )
    cat('\n\n')

}
```

### Expression graphs{.tabset}

#### Broad cell types{.tabset}
```{r act_all_broad, fig.height=5, fig.width=5, results='asis', cache=FALSE}
for (t in levels(conos_dt$type_broad)){
    cat('##### ', t, '\n')
        graph_col_act(graph_all$graph, act_all_broad[t,], graph_all$lay, '')
    cat('\n\n')
}
```

#### Lesion types{.tabset}
```{r act_all_lesion, fig.height=5, fig.width=5, results='asis', cache=FALSE}
for (t in levels(conos_dt$lesion_type)){
    cat('##### ', t, '\n')
        graph_col_act(graph_all$graph, act_all_lesion[t,], graph_all$lay, '')
    cat('\n\n')
}
```

### Expression heatmaps{.tabset}

#### Broad cell types
```{r hm_all_broad, fig.height=35, fig.width=5, results='asis', cache=FALSE}
col_dt = data.table(
    ID    = rownames(act_all_broad), 
    group = rownames(act_all_broad),
    color = broad_cols[rownames(act_all_broad)]) %>%
    setkey(ID)
hm = hm_col_act(
    mtx     = act_all_broad, 
    comm_dt = graph_all$comm_dt,
    col_dt  = col_dt
) 
draw(hm)
```

#### Lesion types
```{r hm_all_lesion, fig.height=35, fig.width=5, results='asis', cache=FALSE}
col_dt = data.table(
    ID    = rownames(act_all_lesion), 
    group = rownames(act_all_lesion),
    color = lesion_cols[rownames(act_all_lesion)]) %>%
    setkey(ID)
hm = hm_col_act(
    mtx     = act_all_lesion, 
    comm_dt = graph_all$comm_dt,
    col_dt  = col_dt,
    cluster_columns=FALSE
) 
draw(hm)
```

#### Samples
```{r hm_all_samples, fig.height=35, fig.width=30, results='asis', cache=FALSE, eval=FALSE}
act_all_samples = act_colors(t(pb_counts_all[graph_all$comm_dt$GENE,]), colnames(pb))
col_dt = data.table(
    ID    = rownames(t(pb_counts_all)), 
    group = colData(pb)$group_id,
    color = lesion_cols[colData(pb)$group_id]) %>%
    setkey(ID)
hm = hm_col_act(
    mtx     = act_all_samples, 
    comm_dt = graph_all$comm_dt,
    col_dt  = col_dt
) 
draw(hm)
```

### UMAP embeddings
```{r redim_all, fig.height=7, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE, cache=FALSE, eval=FALSE}
# calced 
data_norm    = graph_all$data
data_trans   = graph_all$data_trans
cell_types   = conos_dt$type_broad
lesion_types = conos_dt$lesion_type
# umap calcs
umap_norm  = uwot::umap(data_norm)
umap_trans = uwot::umap(data_trans) 
umap_markers = uwot::umap(data_norm[,unique(graph_all$comm_dt$marker)])
# plots
pl_type   = plot_umap(umap_norm, cell_types, label_cols=broad_cols, title='All genes')
pl_lesion = plot_umap(umap_norm, lesion_types, label_cols=lesion_cols, title='')
lp_type   = plot_umap(umap_trans, cell_types, label_cols=broad_cols, title='Projected')
lp_lesion = plot_umap(umap_trans, lesion_types, label_cols=lesion_cols, title='')
lm_type   = plot_umap(umap_markers, cell_types, label_cols=broad_cols, title='Marker genes')
lm_lesion = plot_umap(umap_markers, lesion_types, label_cols=lesion_cols, title='')
(pl_type + pl_lesion + lp_type + lp_lesion + lm_type + lm_lesion) +
    plot_layout(guides = 'collect', nrow=3)
```

## GGM on the broad types{.tabset}
```{r ggm_broad_graph, fig.height=11, fig.width=11 , collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='asis'}
for(t in levels(conos_dt$type_broad)){
    cat('### ', t, '\n')
        gg = graph_col_comm(

            graph  = graphs[[t]]$graph, 
            lay    = graphs[[t]]$lay, 
            grp    = graphs[[t]]$comm_dt$color, 
            sz     = 5, 
            title  = 'GGM on %s', 
            labels = graphs[[t]]$comm_dt$GENE
        )
    cat('\n\n')

}
```


##


### Expression graphs{.tabset}

#### Fine cell types{.tabset}
```{r act_broad_fine, fig.height=5, fig.width=5, results='asis', cache=FALSE}
for (t in levels(conos_dt$type_broad)){
    cat('##### ', t, '{.tabset}\n')
        for (l in unique(conos_dt[type_broad==t]$conos_merge)){
            cat('###### ', l, '\n')
                graph_col_act(graphs[[t]]$graph, act_broad_fine[[t]][l,], graphs[[t]]$lay, '')
            cat('\n\n')
        }
    cat('\n\n')
}
```


#### Lesion types{.tabset}
```{r act_broad_lesion, fig.height=5, fig.width=5, results='asis', cache=FALSE}
for (t in levels(conos_dt$type_broad)){
    cat('##### ', t, '{.tabset}\n')
        for (l in levels(conos_dt[type_broad==t]$lesion_type)){
            cat('###### ', l, '\n')
                graph_col_act(graphs[[t]]$graph, act_broad_lesion[[t]][l,], graphs[[t]]$lay, '')
            cat('\n\n')
        }
    cat('\n\n')
}
```

##

### Expression heatmaps{.tabset}

#### Fine cell types{.tabset}
```{r hm_broad_fine, fig.height=25, fig.width=5, results='asis', cache=FALSE}
for(t in levels(conos_dt$type_broad)){
    cat('##### ', t, '\n')
        col_dt = data.table(
            ID    = rownames(act_broad_fine[[t]]), 
            group = rownames(act_broad_fine[[t]]),
            color = nice_cols[1:dim(act_broad_fine[[t]])[1]]) %>%
            setkey(ID)
        hm = hm_col_act(
            mtx      = act_broad_fine[[t]], 
            comm_dt  = graphs[[t]]$comm_dt,
            col_dt   = col_dt
        ) 
        draw(hm)
    cat('\n\n')
}

```

#### Lesion types{.tabset}
```{r hm_broad_lesion, fig.height=25, fig.width=5, results='asis', cache=FALSE}
for(t in levels(conos_dt$type_broad)){
    cat('##### ', t, '\n')
        col_dt = data.table(
            ID    = rownames(act_broad_lesion[[t]]), 
            group = rownames(act_broad_lesion[[t]]),
            color = lesion_cols[rownames(act_broad_lesion[[t]])]) %>%
            setkey(ID)
        hm = hm_col_act(
            mtx      = act_broad_lesion[[t]], 
            comm_dt  = graphs[[t]]$comm_dt,
            col_dt   = col_dt,
            cluster_columns=FALSE
        ) 
        draw(hm)
    cat('\n\n')
}
```

#### Samples{.tabset}
```{r hm_broad_samples, fig.height=25, fig.width=20, results='asis', cache=FALSE, eval=FALSE}
for(t in levels(conos_dt$type_broad)){
    cat('##### ', t, '\n')
        act_samples = act_colors(t(pb_counts[[t]][graphs[[t]]$comm_dt$GENE,]), colnames(pb_counts[[t]]))
        col_dt = data.table(
            ID    = rownames(t(pb_counts[[t]])), 
            group = colData(pb[, rownames(t(pb_counts[[t]]))])$group_id,
            color = lesion_cols[colData(pb[, rownames(t(pb_counts[[t]]))])$group_id]) %>%
            setkey(ID)
        hm = hm_col_act(
            mtx     = act_samples, 
            comm_dt = graphs[[t]]$comm_dt,
            col_dt  = col_dt
        ) 
        draw(hm)
    cat('\n\n')
}
```



## GGM on the combined types{.tabset}
```{r ggm_comb_graph, fig.height=11, fig.width=11 , collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='asis'}
for(t in names(comb_types)){
    cat('### ', t, '\n')
        gg = graph_col_comm(
            graph  = graphs_comb[[t]]$graph, 
            lay    = graphs_comb[[t]]$lay, 
            grp    = graphs_comb[[t]]$comm_dt$color, 
            sz     = 5, 
            title  = 'GGM on everything', 
            labels = graphs_comb[[t]]$comm_dt$GENE
        )
    cat('\n\n')

}
```


##


### Expression graphs{.tabset}

#### Fine cell types{.tabset}
```{r act_comb_fine, fig.height=5, fig.width=5, results='asis', cache=FALSE}
for (t in names(comb_types)){
    types = comb_types[[t]]
    cat('##### ', t, '{.tabset}\n')
        for (l in unique(conos_dt[type_broad %in% types]$conos_merge)){
            cat('###### ', l, '\n')
                graph_col_act(graphs_comb[[t]]$graph, act_comb_fine[[t]][l,], graphs_comb[[t]]$lay, '')
            cat('\n\n')
        }
    cat('\n\n')
}
```


#### Lesion types{.tabset}
```{r act_comb_lesion, fig.height=5, fig.width=5, results='asis', cache=FALSE}
for (t in names(comb_types)){
    types = comb_types[[t]]
    cat('##### ', t, '{.tabset}\n')
        for (l in levels(conos_dt[type_broad %in% types]$lesion_type)){
            cat('###### ', l, '\n')
                graph_col_act(graphs_comb[[t]]$graph, act_comb_lesion[[t]][l,], graphs_comb[[t]]$lay, '')
            cat('\n\n')
        }
    cat('\n\n')
}
```

##

### Expression heatmaps{.tabset}

#### Fine cell types{.tabset}
```{r hm_comb_fine, fig.height=25, fig.width=5, results='asis', cache=FALSE}
for(t in names(comb_types)){
    cat('##### ', t, '\n')
        col_dt = data.table(
            ID    = rownames(act_comb_fine[[t]]), 
            group = rownames(act_comb_fine[[t]]),
            color = nice_cols[1:dim(act_comb_fine[[t]])[1]]) %>%
            setkey(ID)
        hm = hm_col_act(
            mtx      = act_comb_fine[[t]], 
            comm_dt  = graphs_comb[[t]]$comm_dt,
            col_dt   = col_dt
        ) 
        draw(hm)
    cat('\n\n')
}

```

#### Lesion types{.tabset}
```{r hm_comb_lesion, fig.height=25, fig.width=5, results='asis', cache=FALSE}
for(t in names(comb_types)){
    cat('##### ', t, '\n')
        col_dt = data.table(
            ID    = rownames(act_comb_lesion[[t]]), 
            group = rownames(act_comb_lesion[[t]]),
            color = lesion_cols[rownames(act_comb_lesion[[t]])]) %>%
            setkey(ID)
        hm = hm_col_act(
            mtx      = act_comb_lesion[[t]], 
            comm_dt  = graphs_comb[[t]]$comm_dt,
            col_dt   = col_dt,
            cluster_columns=FALSE
        ) 
        draw(hm)
    cat('\n\n')
}
```



##

### UMAP embeddings{.tabset}
```{r redim_broad, fig.height=5, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE, cache=FALSE, results='asis', eval=FALSE}
for(t in levels(conos_dt$type_broad)){
    cat('#### ', t, '\n')
        # calced 
        data_norm    = graphs[[t]]$data
        data_trans   = graphs[[t]]$data_trans
        cell_types   = conos_dt[type_broad==t]$conos_merge
        lesion_types = conos_dt[type_broad==t]$lesion_type
        # umap calcs
        umap_norm  = uwot::umap(data_norm)
        umap_trans = uwot::umap(data_trans)  
        umap_markers = uwot::umap(data_norm[,unique(graphs[[t]]$comm_dt$marker)])
        # plots
        pl_type   = plot_umap(umap_norm, cell_types, title='All genes')
        pl_lesion = plot_umap(umap_norm, lesion_types, label_cols=lesion_cols, title='')
        lp_type   = plot_umap(umap_trans, cell_types, title='Projected')
        lp_lesion = plot_umap(umap_trans, lesion_types, label_cols=lesion_cols, title='')
        lm_type   = plot_umap(umap_markers, cell_types, title='Marker genes')
        lm_lesion = plot_umap(umap_markers, lesion_types, label_cols=lesion_cols, title='')
        pl = (pl_type + pl_lesion + lp_type + lp_lesion + lm_type + lm_lesion) +
            plot_layout(guides = 'collect', nrow=3)
        print(pl)
    cat('\n\n')
}
```

## Sample distances based on the psudo-bulk object{.tabset}
```{r redim_samples, fig.height=5, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE, cache=FALSE, results='asis', eval=FALSE}
data_norm = list()
data_trans = list()
lesion_types = pb_meta$lesion_type
status_types = pb_meta$disease_status
for(t in levels(conos_dt$type_broad)){
    cat('### ', t, '\n')
        # calced 
        data_norm[[t]]    = t(pb_counts[[t]][rownames(graphs[[t]]$trans), ])
        data_trans[[t]]   = samples_trans[[t]]
        # umap calcs
        umap_norm  = uwot::umap(data_norm[[t]])
        umap_trans = uwot::umap(data_trans[[t]])  
        umap_markers = uwot::umap(data_norm[[t]][,unique(graphs[[t]]$comm_dt$marker)])
        # plots
        pl_lesion = plot_umap(umap_norm, lesion_types, label_cols=lesion_cols, title='All genes')
        pl_status = plot_umap(umap_norm, status_types, label_cols=disease_cols, title='')
        lp_lesion = plot_umap(umap_trans, lesion_types, label_cols=lesion_cols, title='Projected')
        lp_status = plot_umap(umap_trans, status_types, label_cols=disease_cols, title='')
        lm_lesion = plot_umap(umap_markers, lesion_types, label_cols=lesion_cols, title='Marker genes')
        lm_status = plot_umap(umap_markers, status_types, label_cols=disease_cols, title='')
        pl = (pl_type + pl_lesion + lp_type + lp_lesion + lm_lesion + lm_status) +
            plot_layout(guides = 'collect', nrow=3)
        print(pl)
    cat('\n\n')
}
```
##

### Integration
```{r redim_int, fig.height=5, fig.width=7, message=FALSE, warning=FALSE, paged.print=FALSE, cache=FALSE, results='asis', eval=FALSE}
data_norm  = data_norm %>% purrr::reduce(cbind)
data_trans = data_trans %>% purrr::reduce(cbind)
umap_norm  = uwot::umap(data_norm)
umap_trans = uwot::umap(data_trans) 
umap_markers = uwot::umap(data_norm[,unique(graphs[[t]]$comm_dt$marker)]) 
# plots
pl_lesion = plot_umap(umap_norm, lesion_types, label_cols=lesion_cols, title='All genes')
pl_status = plot_umap(umap_norm, status_types, label_cols=disease_cols, title='')
lp_lesion = plot_umap(umap_trans, lesion_types, label_cols=lesion_cols, title='Projected')
lp_status = plot_umap(umap_trans, status_types, label_cols=disease_cols, title='')
lm_lesion = plot_umap(umap_markers, lesion_types, label_cols=lesion_cols, title='Marker genes')
lm_status = plot_umap(umap_markers, status_types, label_cols=disease_cols, title='')
pl = (pl_type + pl_lesion + lp_type + lp_lesion + lm_lesion + lm_status) +
    plot_layout(guides = 'collect', nrow=3)
print(pl)
```


# Session info
```{r sess, message=FALSE, warning=FALSE, paged.print=FALSE}
sessionInfo()
```