---
title: "Exploring the Spatial Mouse Atlas"
author:
- name: Elyas Heidari
  affiliation: 
  - &IMLS Institute for Molecular Life Sciences, University of Zurich, Switzerland
  - Swiss Institute of Bioinformatics (SIB), University of Zurich, Switzerland
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  workflowr::wflow_html:
    code_folding: hide
    toc: true
    toc_float: true
    number_sections: false
---

# Setup / definitions

## Libraries

```{r setup_knitr, include=FALSE}
library('BiocStyle')
options(bitmapType='cairo')
set.seed(1996)
knitr::opts_chunk$set( autodep=TRUE, cache=FALSE, cache.lazy=FALSE, dev='png' )
knitr::opts_knit$set( root.dir='..' )
# wflow_build(files='analysis/sma01_eda.Rmd', view=F, verbose=T, delete_cache=T)
```

```{r setup_libs, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
suppressMessages(source('code/scGCN/dep.R')) 
```

## Helper functions

```{r setup_helpers, message=FALSE, cache=FALSE}
source('code/sma00_utils.R')
source('code/sma01_eda.R')
```

## Inputs

```{r setup_input}
exprs_f    = 'data/exprs.Rds'
metadata_f = 'data/metadata.Rds'
umap_f     = 'data/E8.5_joint_UMAP_df.Rds'
nngraph_f  = 'data/E8.5_neighbourGraph_1.3.Rds'
```

## Outputs

```{r setup_outputs}
```

# Load inputs

```{r load_inputs}
nngraph  = nngraph_f %>% readRDS 
cells    = names(V(nngraph))
exprs    = exprs_f %>% readRDS %>% 
    as.matrix %>% .[, cells]  %>% t 
metadata = metadata_f %>% readRDS %>% as.data.table %>%
    setnames('uniqueID', 'cell') %>%
    setkey('cell') %>%
    .[cells] %>%
    .[, embryo := paste(embryo, z, sep='_')]
# umap     = umap_f %>% readRDS %>% as.data.table %>%
#     setkey('cell') %>%
#     .[cells] 
embryos       = unique(metadata$embryo)
cells_list   = embryos %>%
    map(~unlist(metadata[embryo == .x, 'cell'])) %>%
    setNames(embryos)
nngraph_list = cells_list %>% 
    map(~induced.subgraph(nngraph, .x))
exprs_list   = cells_list %>% 
    map(~exprs[.x, ])

```

# Processing / calculations

## Spatial graph communities
```{r spgraph_comm, message=FALSE, warning=FALSE, paged.print=FALSE}
nncomm_list  = nngraph_list %>% 
    map(nngraph_comm) %>%
    setNames(NULL) %>%
    unlist
metadata$nncomm = paste(metadata$embryo, nncomm_list[metadata$cell], sep='_') 
meta_list    = cells_list %>%
    map(~metadata[cell %in% .x, ])
```

## GGM on everything

```{r ggm_all_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
#TODO: optimize rho and threshold
graph_all = counts2graph(
    mtx       = exprs,
    rho       = 0.2,
    threshold = 0.02,
    no_hvg    = TRUE
)
act_all_types  = act_colors(graph_all$data, metadata$celltype_mapped_refined)
act_all_comms  = act_colors(graph_all$data, metadata$nncomm)
gc()
```


## GGM on embryo sections (robustness)
```{r ggm_embryo_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
#TODO: optimize rho and threshold
graphs_embryo = list()
ari_embryo    = list()
for(t in unique(metadata$embryo)){
    idx = which(metadata$embryo == t)
    log_counts_type = graph_all$data[idx,] 
    if(sum(colSds(log_counts_type) == 0) > 0)
        log_counts_type %<>% .[, -which(colSds(.) == 0)]
    log_counts_type %<>% as.data.table
    ggm_obj = ggm(log_counts_type, rho=0.2, threshold=0.02)
    gg = ggm_obj$graph
    wi = ggm_obj$model$wi
    isolated  = which(igraph::degree(gg) == 0)
    gg = delete.vertices(gg, isolated)
    wi %<>% .[-isolated, -isolated]
    ggm_obj$graph    = gg
    ggm_obj$model$wi = wi
    graphs_embryo[[t]] = ggm_obj
    comms  = leiden(abs(wi) > 0.02)
    comm_dt  = copy(graph_all$comm_dt) %>% setkey('GENE')
    ref_comm = comm_dt[names(V(gg))]$community
    ari_embryo[[t]] = mclust::adjustedRandIndex(ref_comm, comms)
}
gc()
```


## GGM on cell types (robustness)
```{r ggm_type_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
#TODO: optimize rho and threshold
graphs_type = list()
ari_type    = list()
for(t in levels(factor(metadata$celltype_mapped_refined, ordered=T))){
    idx = which(metadata$celltype_mapped_refined == t)
    if(length(idx) < 1000)
        next
    log_counts_type = graph_all$data[idx,] 
    if(sum(colSds(log_counts_type) == 0) > 0)
        log_counts_type %<>% .[, -which(colSds(.) == 0)]
    log_counts_type %<>% as.data.table
    ggm_obj = ggm(log_counts_type, rho=0.15, threshold=0.015)
    gg = ggm_obj$graph
    wi = ggm_obj$model$wi
    isolated  = which(igraph::degree(gg) == 0)
    gg = delete.vertices(gg, isolated)
    wi %<>% .[-isolated, -isolated]
    ggm_obj$graph    = gg
    ggm_obj$model$wi = wi
    graphs_type[[t]] = ggm_obj
    comms  = leiden(abs(wi) > 0.015)
    comm_dt  = copy(graph_all$comm_dt) %>% setkey('GENE')
    ref_comm = comm_dt[names(V(gg))]$community
    ari_type[[t]] = mclust::adjustedRandIndex(ref_comm, comms)
}
gc()
```

## GGM on spatial communities (robustness)
```{r ggm_nncomm_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
#TODO: optimize rho and threshold
graphs_nncomm = list()
ari_nncomm    = list()
for(t in unique(metadata$nncomm)){
    idx = which(metadata$nncomm == t)
    if(length(idx) < 1000)
        next
    log_counts_type = graph_all$data[idx,] 
    if(sum(colSds(log_counts_type) == 0) > 0)
        log_counts_type %<>% .[, -which(colSds(.) == 0)]
    log_counts_type %<>% as.data.table
    ggm_obj = ggm(log_counts_type, rho=0.15, threshold=0.015)
    gg = ggm_obj$graph
    wi = ggm_obj$model$wi
    isolated  = which(igraph::degree(gg) == 0)
    gg = delete.vertices(gg, isolated)
    wi %<>% .[-isolated, -isolated]
    print(dim(wi))
    ggm_obj$graph    = gg
    ggm_obj$model$wi = wi
    graphs_nncomm[[t]] = ggm_obj
    comms  = leiden(abs(wi) > 0.015)
    comm_dt  = copy(graph_all$comm_dt) %>% setkey('GENE')
    ref_comm = comm_dt[names(V(gg))]$community
    ari_nncomm[[t]] = mclust::adjustedRandIndex(ref_comm, comms)
}
gc()
```

## GGM on embryo sections (functional annotation)
```{r ggm_embryo_full_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
#TODO: optimize rho and threshold
graphs            = list()
act_embryo_type   = list()
act_embryo_comm   = list()
for(t in levels(factor(metadata$embryo, ordered=TRUE))){
    print(t)
    graphs[[t]] = counts2graph(
        mtx       = exprs_list[[t]],
        rho       = 0.2,
        threshold = 0.02,
        no_hvg    = TRUE
    )
    act_embryo_type[[t]] = act_colors(graphs[[t]]$data, meta_list[[t]]$celltype_mapped_refined)
    act_embryo_comm[[t]] = act_colors(graphs[[t]]$data, meta_list[[t]]$nncomm) 
}
gc()
```


## GGM on embryo comms (functional annotation)
```{r ggm_comm_calc, collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='hide'}
#TODO: optimize rho and threshold
graphs            = list()
act_embryo_type   = list()
act_embryo_comm   = list()
for(t in levels(factor(metadata$embryo, ordered=TRUE))){
    print(t)
    graphs[[t]] = counts2graph(
        mtx       = exprs_list[[t]],
        rho       = 0.2,
        threshold = 0.02,
        no_hvg    = TRUE
    )
    act_embryo_type[[t]] = act_colors(graphs[[t]]$data, meta_list[[t]]$celltype_mapped_refined)
    act_embryo_comm[[t]] = act_colors(graphs[[t]]$data, meta_list[[t]]$nncomm) 
}
gc()
```


# Analysis

## Spatial plots{.tabset}
```{r sp_plots, fig.height=5, fig.width=12, results='asis', cache=FALSE}
for(t in levels(factor(metadata$embryo, ordered=TRUE))){
    cat('### ', t, '\n')
        type_p = plot_2d(
            dim_df = meta_list[[t]][, .(x_global_affine, -y_global_affine)],
            labels = meta_list[[t]]$celltype_mapped_refined,
            label_cols = celltype_colours,
            label_title  = 'Cell type'
        )
        comm_p = plot_2d(
            dim_df = meta_list[[t]][, .(x_global_affine, -y_global_affine)],
            labels = meta_list[[t]]$nncomm,
            label_title = 'Spatial neighborhood'
        )     
        pl = (type_p + comm_p) + plot_layout(guides = 'collect')
        print(pl)
    cat('\n\n')

}

```

## GGM on everything 
```{r ggm_all_graph, fig.height=11, fig.width=11 , collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
graph_col_comm(
    graph  = graph_all$graph, 
    lay    = graph_all$lay, 
    grp    = graph_all$comm_dt$color, 
    sz     = 5, 
    title  = 'GGM on everything', 
    labels = graph_all$comm_dt$GENE
)
```


### GGM on embryo sections (robustness){.tabset}
```{r ggm_embryo_rob_graph, fig.height=11, fig.width=11, results='asis' , collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
comm_dt = copy(graph_all$comm_dt) %>% setkey(GENE)
for(t in levels(factor(metadata$embryo, ordered=TRUE))){
    cat('#### ', t, '\n')
        gg = graph_col_comm(
            graph  = graphs_embryo[[t]]$graph, 
            lay    = layout_nicely(graphs_embryo[[t]]$graph), 
            grp    = comm_dt[names(V(graphs_embryo[[t]]$graph))]$color, 
            sz     = 5, 
            title  = sprintf('GGM on %s, ARI = %s', t, ari_embryo[[t]]), 
            labels = comm_dt[names(V(graphs_embryo[[t]]$graph))]$GENE
        )
    cat('\n\n')

}
```

### GGM on cell types (robustness){.tabset}
```{r ggm_type_rob_graph, fig.height=11, fig.width=11, results='asis' , collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
comm_dt = copy(graph_all$comm_dt) %>% setkey(GENE)
for(t in levels(factor(names(graphs_type), ordered=TRUE))){
    cat('#### ', t, '\n')
        gg = graph_col_comm(
            graph  = graphs_type[[t]]$graph, 
            lay    = layout_nicely(graphs_type[[t]]$graph), 
            grp    = comm_dt[names(V(graphs_type[[t]]$graph))]$color, 
            sz     = 5, 
            title  = sprintf('GGM on %s, ARI = %s', t, ari_type[[t]]), 
            labels = comm_dt[names(V(graphs_type[[t]]$graph))]$GENE
        )
    cat('\n\n')

}
```

### GGM on spatial communities (robustness){.tabset}
```{r ggm_nncomm_rob_graph, fig.height=11, fig.width=11, results='asis' , collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
comm_dt = copy(graph_all$comm_dt) %>% setkey(GENE)
for(t in levels(factor(names(graphs_nncomm), ordered=TRUE))){
    cat('#### ', t, '\n')
        gg = graph_col_comm(
            graph  = graphs_nncomm[[t]]$graph, 
            lay    = layout_nicely(graphs_nncomm[[t]]$graph), 
            grp    = comm_dt[names(V(graphs_nncomm[[t]]$graph))]$color, 
            sz     = 5, 
            title  = sprintf('GGM on %s', t), 
            labels = comm_dt[names(V(graphs_nncomm[[t]]$graph))]$GENE
        )
    cat('\n\n')

}
```

### Expression graphs{.tabset}

#### On cell types{.tabset}
```{r act_all_types, fig.height=5, fig.width=5, results='asis', cache=FALSE}
for (t in unique(metadata$celltype_mapped_refined)){
    cat('##### ', t, '\n')
        graph_col_act(graph_all$graph, act_all_types[t,], graph_all$lay, '')
    cat('\n\n')
}
```

#### On spatial graph communities {.tabset}
```{r act_all_comms, fig.height=5, fig.width=5, results='asis', cache=FALSE}
for (t in unique(metadata$nncomm)){
    cat('##### ', t, '\n')
        graph_col_act(graph_all$graph, act_all_comms[t,], graph_all$lay, '')
    cat('\n\n')
}
```

### Expression heatmap{.tabset} 

#### On cell types

```{r hm_all_type, fig.height=25, fig.width=10, results='asis', cache=FALSE}
col_dt = data.table(
    ID    = rownames(act_all_types), 
    group = rownames(act_all_types),
    color = celltype_colours[rownames(act_all_types)]) %>%
    setkey(ID)
hm = hm_col_act(
    mtx     = act_all_types, 
    comm_dt = graph_all$comm_dt,
    col_dt  = col_dt
) 
draw(hm)
```

#### On spatial graph communities
```{r hm_all_comm, fig.height=25, fig.width=15, results='asis', cache=FALSE}
col_dt = data.table(
    ID    = rownames(act_all_comms), 
    group = rownames(act_all_comms),
    color = nice_cols[as.numeric(factor(rownames(act_all_comms), ordered=T))]) %>%
    setkey(ID)
hm = hm_col_act(
    mtx     = act_all_comms, 
    comm_dt = graph_all$comm_dt,
    col_dt  = col_dt
) 
draw(hm)
```


## GGM on the embryo sections (functional annotation){.tabset}
```{r ggm_embryo_graph, fig.height=11, fig.width=11 , collapse=FALSE, message=FALSE, warning=FALSE, cache=FALSE, results='asis'}
for(t in levels(factor(metadata$embryo, ordered=TRUE))){
    cat('### ', t, '\n')
        gg = graph_col_comm(
            graph  = graphs[[t]]$graph, 
            lay    = graphs[[t]]$lay, 
            grp    = graphs[[t]]$comm_dt$color, 
            sz     = 5, 
            title  = 'GGM on %s', 
            labels = graphs[[t]]$comm_dt$GENE
        )
    cat('\n\n')

}
```


##


### Expression graphs{.tabset}

#### On cell types{.tabset}
```{r act_embryo_type, fig.height=5, fig.width=5, results='asis', cache=FALSE}
for (t in levels(factor(metadata$embryo, ordered=TRUE))){
    cat('##### ', t, '{.tabset}\n')
        for (l in unique(meta_list[[t]]$celltype_mapped_refined)){
            cat('###### ', l, '\n')
                graph_col_act(graphs[[t]]$graph, act_embryo_type[[t]][l,], graphs[[t]]$lay, '')
            cat('\n\n')
        }
    cat('\n\n')
}
```


#### On spatial graph communities{.tabset}
```{r act_embryo_comm, fig.height=5, fig.width=5, results='asis', cache=FALSE}
for (t in levels(factor(metadata$embryo, ordered=TRUE))){
    cat('##### ', t, '{.tabset}\n')
        for (l in unique(meta_list[[t]]$nncomm)){
            cat('###### ', l, '\n')
                graph_col_act(graphs[[t]]$graph, act_embryo_comm[[t]][l,], graphs[[t]]$lay, '')
            cat('\n\n')
        }
    cat('\n\n')
}
```

##

### Expression heatmaps{.tabset}

#### On cell types{.tabset}
```{r hm_embryo_type, fig.height=25, fig.width=8, results='asis', cache=FALSE}
for(t in levels(factor(metadata$embryo, ordered=TRUE))){
    cat('##### ', t, '\n')
        col_dt = data.table(
            ID    = rownames(act_embryo_type[[t]]), 
            group = rownames(act_embryo_type[[t]]),
            color = celltype_colours[rownames(act_embryo_type[[t]])]) %>%
            setkey(ID)
        hm = hm_col_act(
            mtx      = act_embryo_type[[t]], 
            comm_dt  = graphs[[t]]$comm_dt,
            col_dt   = col_dt
        ) 
        draw(hm)
    cat('\n\n')
}

```

#### On spatial graph communities{.tabset}
```{r hm_embryo_comm, fig.height=25, fig.width=5, results='asis', cache=FALSE}
for(t in levels(factor(metadata$embryo, ordered=TRUE))){
    cat('##### ', t, '\n')
        col_dt = data.table(
            ID    = rownames(act_embryo_comm[[t]]), 
            group = rownames(act_embryo_comm[[t]]),
            color = nice_cols[as.numeric(factor(rownames(act_embryo_comm[[t]]), ordered=T))]) %>%
            setkey(ID)
        hm = hm_col_act(
            mtx      = act_embryo_comm[[t]], 
            comm_dt  = graphs[[t]]$comm_dt,
            col_dt   = col_dt,
            cluster_columns=TRUE
        ) 
        draw(hm)
    cat('\n\n')
}
```


# Outputs

```{r save_outputs}
```

```{r session_info, include=TRUE, echo=TRUE, results='markup'}
devtools::session_info()
```
